Empirical normative theories

### 2020-06-25 simple.tex
Rm old successors and prop about finite S
Clean falsifiability VS convincing

\section{Additive single-answer empirical theories}
\subsection{Definition}
\commentOC{I leave the definition as is, but for the moment I only consider ${\gleadsto}$, a relation that does not depend on $i$; and I do not use $q(i)$.}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, \alldargs, \dbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq (\allargs × \Phi) \cap (\alldargs × \Phi)$, 
${\gbeats} \subseteq \allargs × \allargs$,
${\dbeats} \subseteq \alldargs × \allargs$ and is injective (thus $\forall \ar \in \allargs: {\dbeatsinv}(\ar)$ is a singleton or empty). 
Define $\df \subseteq \allargs × \alldargs$ as $\df = {\dbeatsinv} \circ {\gbeatsinv}$. 
The symbol $+$ in an empirical theory stands for a function $+: \bigcup_{\ar \in \allargs}(\set{s} × \df(\ar)) → \allargs$.
As usual, $\ar + \ard$ means $+(\ar, \ard)$. 
Define $\dfp \subseteq \allargs × \allargs$ as, $\forall \ar \in \allargs$: $\dfp(\ar) = \ar + \df(\ar) = \set{\ar + \ard \suchthat \ard \in \df(\ar)}$.
An empirical theory must satisfy
$\forall \ar \in {\gleadsto}(\allargs): \exists k \in \N^* \suchthat \dfp^k(\ar) = \emptyset$.

@@

\begin{proposition}
	$\forall i \in I, \phi \in \Phi, \ar_0 \in \allargs$:
\begin{equation}
	\label{eq:st-in-terms-of-e}
	\ar_0 \nileadstosti \phi ⇔ [\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatsi \ar_0].
\end{equation}
\end{proposition}
\begin{proof}
	By definition of ${\ileadstosti}$, $\ar_0 \ileadstosti \phi ⇔ \forall \hist \in \allhist \suchthat \ar_0 \in \histend: \hist \ileadstosti \phi$. 
Equivalently, $\ar_0 \ileadstosti \phi ⇔ \forall \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend]: \hist \ileadstosti \phi$. 
Still equivalently, $\ar_0 \nileadstosti \phi ⇔ \exists \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend] \land \hist \nileadstosti \phi$. 
This can also be written: $\ar_0 \nileadstosti \phi ⇔ \exists \hist \in \allhist, \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend \land \hist \nileadstosti \phi$. By definition of ${\ibeatsi}$, this is the proposition to be proven.
\end{proof}

Given $i \in I, \phi \in {\gleadsto}(\allargs), \ar \in \allargs$, define the \emph{$\phi$-defensors of $\ar$ required for $i$}, $d^{i, \phi}(\ar) \subseteq \alldargs$, as $d^{i, \phi}(\ar) = [{\dbeatsinv} \circ ({\gbeatsinv} \cap {\ibeatsiinv})](\ar)$.
%Given $i \in I, \phi \in {\gleadsto}(\allargs), \ar \in \allargs, \ard \in \df(\ar)$, define the \emph{successors} of $(\ar, \ard)$ as $\sigma^{i, \phi}(\ar, \ard) = \set{\ar + \ard} × d^{i, \phi}(\ar + \ard)$. 
Given $i \in I, \phi \in {\gleadsto}(\allargs)$, define $\gargs[i, \phi] \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\phi)$ under $\dfp$: $\gargs[i, \phi] = \cup_{k \in \N} \dfp^k({\gleadstoinv}(\phi))$, where $\dfp^0$ is the identity function.

%\begin{proposition}
%	An empirical theory has only finite sets $\gargs[i, \phi]$, and an upper bound on their size exists that is independent of ${\ileadsto}$.
%\end{proposition}
%\begin{proof}
%	Given $i, \phi$. Observe that $\forall \ard: d^{i, \phi}(\ard) \subseteq {\gbeatsmt}(\ard)$. Thus, $d^{i, \phi}$ is finite. Also, the transitive closure through $\sigma^i$ changes a finite number of time, otherwise it is possible to build an infinite chain in ${\gbeatsmt}$.
%\end{proof}

$\ar_0$ $i$-defends $\phi$ when ignoring $\args_1$ iff $\forall \hist \in \allhist: [\args_1 \cap \hist = \emptyset] ⇒ (\hist, \ar_0) \ileadstosts \phi$.
$\ar_0$ $i$-defends $\phi$ under protection of $\args_2$ iff $\ar_0$ $i$-defends $\phi$ when ignoring $g(\ar_0)(\args_2)$.
$\ar_0$ $i$-defends $\phi$ under d-protection of $\argsd_2$ iff $\ar_0$ $i$-defends $\phi$ when ignoring ${\gbeats}(\argsd_2) \cap {\gbeatsinv}(\ar_0)$.

\begin{definition}[Conv. single-try]
	$\forall i \in I, \phi \in {\gleadsto}(\allargs), (\phi, \ar, \ard) \in \gargs[i, \phi], \ar_0 = \ar + \ard: \ar_0$ $i$-defends $\phi$ under d-protection of $d^{i, \phi}(\ard)$.
\end{definition}

1. Checkpath: this is completely general.
2. Only using model that counter-argues: requires a sort of path independence. (arguments can be condensed: s5, s4, s1 => s1big)

@@

A falsification instance of an empirical theory is a tuple $(i, \phi, \ar, \ard, \hist, \ar_1)$ such that $(\phi, \ar, \ard) \in \gargs[i, \phi]$ and, defining $\ar_0 = \ar + \ard$: $[{\gbeatsinv}(\ard) \cap \hist = \emptyset] \land [\hist \text{ ends with } \set{\ar_1, \ar_0}] \land \ar_1 \ibeatse[!\phi] \ar_0]$ (the theory says that this history should defend $\phi$ under prot of something and that is not in the history).

\begin{theorem}
	An empirical theory is conv.\ single-try iff it has no falsification instances.
\end{theorem}
\begin{proof}
	Given $i, \phi, (\phi, \ar, \ard) \in \gargs[i, \phi]$. Define $\args_1 = {\gbeats}(d^{i, \phi}(\ard)) \cap {\gbeatsinv}(\ar)$. Now $\ar$ $i$-defends $\phi$ ignoring $\args_1$ is equivalent to: $\forall \hist: \hist \cap \args_1 = \emptyset ⇒ (\hist, \ar) \ileadstosts \phi$. Negation: $\exists \hist \suchthat \hist \cap \args_1 = \emptyset \land (\hist, \ar) \nileadstosts \phi$. TODO prove that equivalent to no falsification.
\end{proof}

\subsection{Vrac from here}
An empirical theory is defined only given some normative theory. In the following, any reference to an empirical theory should always be understood as an empirical theory for some (sometimes implicitly defined) normative theory.

>>>>

\section{Additive single-answer empirical theories}
\subsection{Definition}
\commentOC{I leave the definition as is, but for the moment I only consider ${\gleadsto}$, a relation that does not depend on $i$; and I do not use $q(i)$.}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, \alldargs, \dbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq (\allargs × \Phi) \cap (\alldargs × \Phi)$, 
${\gbeats} \subseteq \allargs × \allargs$,
${\dbeats} \subseteq \alldargs × \allargs$ and is injective and its range equals ${\gbeatsinv}(\allargs)$, the domain of ${\gbeats}$ (thus $\forall \ar_1 \suchthat [\exists \ar_0 \suchthat \ar_1 \gbeats \ar_0]$: ${\dbeatsinv}(\ar_1)$ is a singleton). 
Define $\df \subseteq \allargs × \alldargs$ as $\df = {\dbeatsinv} \circ {\gbeatsinv}$. 
The symbol $+$ in an empirical theory stands for a function $+: \bigcup_{\ar \in \allargs}(\set{s} × \df(\ar)) → \allargs$.
As usual, $\ar + \ard$ means $+(\ar, \ard)$. 
Define $\dfp \subseteq \allargs × \allargs$ as, $\forall \ar \in \allargs$: $\dfp(\ar) = \ar + \df(\ar) = \set{\ar + \ard \suchthat \ard \in \df(\ar)}$.
An empirical theory must satisfy
$\forall \ar \in {\gleadsto}(\allargs): \exists k \in \N^* \suchthat \dfp^k(\ar) = \emptyset$.

@@

Given $i \in I, \phi \in {\gleadsto}(\allargs), \ar \in \allargs$, define the \emph{$\phi$-defensors of $\ar$ required for $i$}, $d^{i, \phi}(\ar) \subseteq \alldargs$, as $d^{i, \phi}(\ar) = [{\dbeatsinv} \circ ({\gbeatsinv} \cap {\ibeatsiinv})](\ar)$. Define $d^{i, \phi}_+$ as $d^{i, \phi}_+(\ar) = \ar + d^{i, \phi}(\ar)$.
Given $i \in I, \phi \in {\gleadsto}(\allargs)$, define $\gargs[i, \phi] \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\phi)$ under $d^{i, \phi}_+$: $\gargs[i, \phi] = \cup_{k \in \N} (d^{i, \phi}_+)^k({\gleadstoinv}(\phi))$, where $(d^{i, \phi}_+)^0$ is the identity function.

Given $\hist \in \allhist$, $\hist(\intvl{1, l(\hist)})$, where $l(\hist) \in \N$ is the length of the sequence $\hist$, denotes the range of $\hist$, that is, all arguments contained in the sequence $\hist$. Abusing notation, given $\args_1 \subseteq \allargs$, I write $\args_1 \cap \hist$ to designate $\args_1 \cap \hist(\intvl{1, l(\hist)})$, therefore using simply $\hist$ in that expression to denote the \emph{range} of $\hist$.

$\ar$ $i$-defends $\phi$ iff $\forall \hist \in \allhist \suchthat \left[\ar \in \histend \land {\gbeatsinv}(\ar) \cap \hist = \emptyset\right]: \hist \ileadstosti \phi$.

\begin{definition}[Convincingness]
	$\forall i \in I, \phi \in {\gleadsto}(\allargs), \ar \in \gargs[i, \phi]: \ar$ $i$-defends $\phi$.
\end{definition}

A \emph{falsification instance} of an empirical theory is a tuple $(i, \phi, \hist) \in I × {\gleadsto}(\allargs) × \allhist$ such that $\exists \ar \in \histend \cap \gargs[i, \phi] \suchthat {\gbeatsinv}(\ar) \cap \hist = \emptyset \land \hist \nileadstosti \phi$.

\begin{theorem}
	An empirical theory is convincing iff it has no falsification instances.
\end{theorem}
\begin{proof}
	For all $i \in I, \phi \in {\gleadsto}(\allargs)$: $\exists \ar \in \gargs[i, \phi]$ that does not $i$-defend $\phi$ iff $\exists \ar \in \gargs[i, \phi], \hist \in \allhist \suchthat \left[\ar \in \histend \land {\gbeatsinv}(\ar) \cap \hist = \emptyset \land \hist \nileadstosti \phi\right]$ iff $\exists \hist \in \allhist$ such that $(i, \phi, \hist)$ is a falsification instance.

	For all $i \in I, \phi \in {\gleadsto}(\allargs)$, the following equivalences hold. $\forall \ar \in \gargs[i, \phi]$, $\ar$ $i$-defends $\phi$ iff $\forall \ar \in \gargs[i, \phi], \hist \suchthat \left[\ar \in \histend \land {\gbeatsinv}(\ar) \cap \hist = \emptyset\right]$, $\hist \ileadstosti \phi$ iff $\forall \hist \suchthat \left[\ar \in \histend \land {\gbeatsinv}(\ar) \cap \hist = \emptyset\right]$, $(i, \phi, \hist)$ is not a falsification instance.
\end{proof}

\begin{proposition}
	$\forall i \in I, \phi \in \Phi, \ar_0 \in \allargs$:
\begin{equation}
	\label{eq:st-in-terms-of-e}
	\ar_0 \nileadstosti \phi ⇔ [\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatsi \ar_0].
\end{equation}
\end{proposition}
\begin{proof}
	By definition of ${\ileadstosti}$, $\ar_0 \ileadstosti \phi ⇔ \forall \hist \in \allhist \suchthat \ar_0 \in \histend: \hist \ileadstosti \phi$. 
Equivalently, $\ar_0 \ileadstosti \phi ⇔ \forall \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend]: \hist \ileadstosti \phi$. 
Still equivalently, $\ar_0 \nileadstosti \phi ⇔ \exists \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend] \land \hist \nileadstosti \phi$. 
It can also be written $\ar_0 \nileadstosti \phi ⇔ \exists \hist \in \allhist, \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend \land \hist \nileadstosti \phi$. By definition of ${\ibeatsi}$, this is the proposition to be proven.
\end{proof}

\subsection{Vrac from here}
1. Checkpath: this is completely general.
2. Only using model that counter-argues: requires a sort of path independence. (arguments can be condensed: s5, s4, s1 => s1big)

@@

An empirical theory is defined only given some normative theory. In the following, any reference to an empirical theory should always be understood as an empirical theory for some (sometimes implicitly defined) normative theory.
